{\normalsize \textbf{Vagueness and imprecise imitation in signalling games}}

### 2016-06-05 paper/version_final/paper.tex
added acknowledgements, deleted \emph, swapped double- for single quotes

\begin{abstract}
  Signalling games are popular models for studying the evolution of meaning, but typical
  approaches do not incorporate vagueness as a feature of successful signalling.  Complementing
  recent like-minded models, we describe an aggregate population-level dynamic that describes a
  process of imitation of successful behaviour under imprecise perception and realisation of
  similar stimuli. Applying this new dynamic to a generalisation of Lewis' signalling games, we
  show that stochastic imprecision leads to vague, yet by-and-large efficient signal use, and,
  moreover, that it unifies evolutionary outcomes and helps avoid sub-optimal
  categorisation. The upshot of this is that we see \emph{as-if}-generalisation at an aggregate
  level, without agents actually generalising.
\end{abstract}

@@

Formally, a sim-max game consists of a set of states $\States$, a set of messages $\Messgs$
typically with much fewer messages than states, a probability distribution
$\Pr \in \Delta(\States)$ such that $\Pr(\state)$ gives the probability that state $\state$
occurs, a similarity metric\footnote{A metric is a function of distance between any two points
  in a given space. It should satisfy certain axioms that ensure behaviour that one would
  intuitively expect from the words ``similarity'' and ``distance'' alone, but these details do
  not matter for the purposes of this paper. The assumption that state similarity forms a
  metric is conceptually loaded, but we follow the literature here and conceive of it as a
  first and pragmatic simplification, possibly to be dispensed with later.}
  % : similarity should always be non-negative, zero only for the distance
  % between an element and itself, symmetric, and satisfy the triangle inequality that
  % $\similarity(x,z) \le \similarity(x,y) + \similarity(y,z)$
on states $\similarity \mycolon \States \times \States \rightarrow \mathds{R}$ such that
$\similarity(\mystate{1}, \mystate{2})$ is the (physical) similarity between $\mystate{1}$ and
$\mystate{2}$, and a utility function
$\Utils \mycolon \States \times \States \rightarrow \mathds{R}$ such that
$\Utils(\mystate{1}, \mystate{2})$ is the payoff for sender and receiver for a play with actual
state $\mystate{1}$ and receiver interpretation $\mystate{2}$. We identify the receiver's acts
with the states of the world, so that the game is one of guessing the actual state, so to
speak. For the modelling purposes of this paper, we make the simplifying assumption that
$\States$ is a set of points in Euclidean space, whose closeness to each other tracks physical
similarity. Perceived similarity, where it is necessary, would be a monotonic function of
physical similarity. Likewise, the utility function should be a monotonically increasing
function of physical similarity.\footnote{Section~\ref{sec:setting-stage} motivates particular
  choices of similarity and utility functions that we will explore in more detail.}
Non-probabilistic sender behaviour can be represented by a pure strategy
$\Spure \in \Messgs^\States$, which deterministically defines which message would be used for
each state. Similarly, a pure receiver strategy is a function $\Rpure \in \States^\Messgs$.

@@

It is intuitive to think that linguistic and conceptual categories are orderly in such a
manner.  For example, if you consider two people `tall', one with a height of 2m and another with a
height of 2.2m, it would be difficult to defend not considering a person with a height
of 2.1m `tall' as well. Another example where this intuition is additionally supported by empirical data is that of colour categorisation. The World Color Survey
project~(\citealtbjps{CookKay2005:The-World-Color}; \citealtbjps{WCS}) collected colour naming data for 110 unwritten
languages of 45 language families.  For the great majority of these languages a pattern can be
observed: basic colour terms are by and large convex~(\citealtbjps{Regier07}; \citealtbjps{Jager2010}). It
is premature to argue that these observations can be extended to \emph{all} cases of
categorisation. However, for the cases where it does apply, the result of
\citetbjps{JagerMetzger2011:Voronoi-Languag} is interesting because it demonstrates that signalling
can impose this kind of orderly categories on a metric space without that being the ulterior
purpose of it all.%
\footnote{For the concrete case of colour categorisation, see also
  \citetbjps{JagerRooijvan-Rooij2007:Language-Struct} and \citetbjps{Correia2015}}
Finally, these considerations are in line with a prominent
school of thought in comparative linguistics which also assumes that more abstract conceptual domains
(e.g., spatial-topological relations, temporal reference, or the meanings expressible by
indefinite pronouns) are preferably carved up by the languages of the world in such a way
that meanings are connected regions on a so-called \emph{semantic map}
(e.g.~\citealtbjps{Croft2003:Typology-and-Un}; \citealtbjps{Haspelmath2003:The-geometry-of}; \citealtbjps{LevinsonMeira2003:Natural-concept}).

There are at least two potential ways of interpreting the signalling setup.
Sender and receiver can be distinct
entities, whose purpose is to communicate effectively about the actual
state. In that case, evolving Voronoi languages would explain why
\emph{linguistic categories} are well-behaved and orderly in the way
they appear to be. More abstractly, sender and receiver can also be
thought of as distinct modules in a single system, where the first
module must discretise the information it is fed by selecting a small
sample of, suggestively, category labels. These are passed to a second
module that tries to decode the original information. In this case,
evolving Voronoi languages would explain why \emph{conceptual
  categories} are well-behaved and orderly in the way that they appear
to be.
Seen in this light, sim-max games may provide a foundation to approaches in cognitive semantics
that rely on the notion of conceptual spaces.
\citetbjpspost{pp.~70--77}{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued that
natural categories are convex regions in conceptual space. If the conceptual space has a
suitable metric, convex categories can be derived from a set of prototypes. The category
corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
other. In this way, \citeauthor{Gardenfors2000:Conceptual-Spac} argues, an efficient categorisation
system can be obtained: storing the prototypes lets us recover the categories without having to
store each category's extension. However, what is left unexplained is where the prototypes
come from, and why we would not see just any distribution of prototypes as an equally efficient
classification system. This is where sim-max games can contribute a principled approach to
deriving, in an independent way, not only convex categories but also prototypical exemplars
belonging to them.  These ideas, and more, are developed further in
(\citealtbjps{Jager2007:The-Evolution-o}; \citealtbjps{JagerRooijvan-Rooij2007:Language-Struct}; \citealtbjps{JagerMetzger2011:Voronoi-Languag}; \citealtbjps{OConnor2014-OCOEPC}),
among others.

@@

\citetbjps{DouvenDecock2011:Vagueness:-A-Co} show that
\citeauthor{Gardenfors2000:Conceptual-Spac}'s conceptual spaces
approach can be extended to account for the existence of borderline
cases. From the assumption that prototypes are extended, yet convex
regions in conceptual space, a construction algorithm is available
that yields ``collated Voronoi diagrams'' with thick boundaries
representing borderline
regions. \citetbjps{DecockDouven2012:What-is-Graded-} show further how it
is possible to arrive at a gradient transition between categories, by
weighing in the distance of different borderline cases to various
prototypical regions. This accounts for the first of the two
desiderata mentioned above, but still assumes that crisp prototype
regions must be given.

@@

Our conceptual starting point is the widely shared conviction that language is shaped by at
least two forces, which may, on occasion, pull in opposite direction. On the one hand, there is
functional pressure towards efficient communication. On the other hand, there is systematic
error, noise or imprecision in the transmission of linguistic behaviour, knowledge or traits. As
an example for the latter, consider a child learning syntactic rules from a parent
generation. The child must infer these unobservable rules from observable speech. Inductive
biases may influence which syntactic rules are likely to be inferred from (finite) parental
input. Over the course of many generations, the effects of such biases can, in a manner of
speaking, ``accumulate'' and lead to surprising results, such as the evolution of compositional
form-meaning mappings or the use of regular recursive syntactic structure. This can happen when
the effects of transmission biases are isolated, as in iterated learning models
(\citealtbjps{KirbyHurford2002:The-Emergence-o}; \citealtbjps{SmithKirby2003:Iterated-Learni}; \citealtbjps{GriffithsKalish2007:Language-Evolut}; \citealtbjps{KirbyGriffith2014:Iterated-Learni}),
or when they interact with functional pressure towards efficient communication, e.g., as
formalised in the replicator mutator dynamic
(e.g.~\citealtbjps{NowakPlotkin2000:The-Evolution-o}; \citealtbjps{NowakKomarova2001:Evolution-of-Un}).

@@

Fix a sim-max game with finite states $\States$ and messages $\Messgs$. As usual, we assume
that the receiver chooses states in $\States$ in response to messages. Let
$\Pr(\cdot) \in \Delta(\States)$ be the prior distribution over states and
$\Util \mycolon \States \times \States \rightarrow \mathds{R}$ the utility function shared by
senders and receivers in the population. A \emph{behavioural strategy} is a function that maps
an agent's choice points to a probability distribution over available
choices.\footnote{\label{fn:mixedStrats}Our focus is on behavioural strategies, not \emph{mixed
    strategies}, i.e., probability distributions over functions from each choice point to an
  act, such as $s \in \Delta(\Messgs^{\States})$.  Dynamics on behavioural strategies assume
  that agents can adapt their behaviour locally, i.e., independently at each choice point. Our
  focus on behavioural strategies greatly reduces the complexity of the dynamic and simplifies
  numerical simulations. But it also seems the more plausible choice for imitation-based update
  protocols of the kind we consider here: agents only observe how, on some occasion, some other
  agent behaved in \emph{one} particular situation, not how that agent would behave in
  \emph{all} relevant choice situations; they imitate the use of a single word, so to speak,
  not a whole lexicon. (See \citealtbjps{Cressman2003:Evolutionary-Dy} for more on the difference
  between dynamics on mixed or behavioural strategies.) It is in this respect that the cultural
  evolutionary dynamic introduced here differs most visibly from the replicator mutator dynamic
  (e.g.~\citealtbjps{NowakPlotkin2000:The-Evolution-o}; \citealtbjps{NowakKomarova2001:Evolution-of-Un}), which
  operates on mixed strategies and is motivated by assumptions of (asexual) biological
  inheritance with transmission infidelity.} The sender's behavioural strategies are
functions $\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$
to a probability of each message $\messg \in \Messgs$ being sent in $\state$; the receiver's
are functions $\Rstrat \in \Delta(\States)^\Messgs$, thus mapping each message
$\messg \in \Messgs$ to a probability of each interpretation $\state \in \States$ being chosen
in response to $\messg$. Although behavioural strategies are probabilistic, evolutionary
modelling usually imagines that every individual agent has a non-probabilistic
strategy. Behavioural strategies then capture average population behaviour. Assuming a virtually
infinite population, the number $\Sstrat(\messg \probbar \state)$, for instance, is then the
probability that a randomly sampled sender would send message $\messg$ if the actual state was
$\state$. Similarly, $\Rstrat(\state \probbar \messg)$ is then the probability with which a
randomly sampled receiver interprets $\messg$ as $\state$. A play of a single evolutionary game
with behavioural strategies is illustrated in Figure~\ref{fig:sim-max-illustration}.

@@

The \emph{expected utility} of choices at each choice point is:\footnote{The notation
  $\propto$, for ``proportional to'', that is used here and hereafter means that the right-hand
  side might still need to be normalised so as to have probabilities sum to one. In general,
  writing $P(x) \propto f(x)$ for any function $f \mycolon X \rightarrow \mathds{R}$ is
  shorthand for $P(x) = \frac{f(x)}{\sum_{x' \in X} f(x')}$. }
\begin{align*}
  \EU(\messg, \state, \Rstrat) & = \sum_{\state \in \States}
  \Rstrat(\state' \probbar \messg) \myts U(\state, \state') \\
  \EU(\state', \messg, \Sstrat) & = \sum_{\state \in
    \States} P(\state \probbar \messg)  \myts
  U(\state, \state')\,, \text{\ \  where } &
P(\state \probbar \messg) & \propto \Pr(\state) \myts \Sstrat(\messg \probbar \state) \,.
\end{align*}
Expected utilities are sums, over all possible concrete outcomes, of the utilities of these
outcomes, weighted by how likely they occur. For instance, the expected utility for the sender
of choosing message $\messg$, given state $\state$ and a receiver strategy $\Rstrat$ is the
sum, for each interpretation $\state'$, of the probability that some random receiver will
choose $\state'$ given $\messg$ times the actual utility of pair $\state$ and
$\state'$.

The discrete-time replicator dynamic tracks changes in frequency of choices in the population
as proportional to their expected utilities
\citepbjpspre{e.g.}{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The formulation given in
  Equation~(\ref{eq:RDDiscrete}) is adequate only for cases like the one we will be looking at
  in Section~\ref{sec:exploring-rdd}, where utilities are always non-negative and expected
  utilities are always positive.}
\begin{align}
  \label{eq:RDDiscrete}
  \Sstrat'(\messg \probbar \state) & \propto \Sstrat(\messg \probbar \state) \myts
    \EU(\messg, \state,\Rstrat)\,, & \Rstrat'(\state \probbar \messg) & \propto \Rstrat(\state \probbar \messg) \myts
    \EU(\state, \messg,\Sstrat)\,.   
\end{align}
For a given choice point, say a state $\state$, the probability of seeing $\messg$ played by an
average agent in the population \emph{after} the update is proportional to the probability of
seeing it \emph{before} the update, which is $\Sstrat(\messg \probbar \state)$, times the
expected utility of $\messg$ at state $\state$. Intuitively put, frequencies of choices change
by a gradient of current frequency and a measure of how good they are.

The replicator dynamic is an abstract population-level dynamic that describes the mean expected
change of behavioural dispositions in a population of signallers. There are several ways of
deriving the replicator dynamic from agent-level processes of behavioural adaptation. We focus
here on one of the simplest: \emph{imitation of success}
\citepbjpspre{see}{Sandholm2010:Population-Game}. The intuitive idea is the following (see
Appendix~\ref{sec:cond-imit-succ} for a derivation of the standard replicator dynamic from this
update scheme).  Every now and then a random agent gets a chance to alter his behaviour for one
of his choice points. Say, this is a sender who gets to ``reconsider'' his choice of message
for state $\state$. The revising agent then observes what some random agent does at $\state$,
say $\messg$, and will henceforth play $\messg$ in $\state$ with a probability given by the
expected utility of $\messg$ for state $\state$.

@@

How does a tendency to confuse similar states interact with selective pressure towards more
efficient signalling strategies under the imprecise conditional imitation dynamic? If confusion
probabilities are moderate and regular, in that they track similarity of states, we might
expect a regularising effect also on evolving signalling strategies. Indeed, we hypothesise that
confusion of states can give rise to population-level aggregate behaviour that looks \emph{as
  if} information about what is good for one state percolates also to similar states. In other
words, imprecise imitation may make signalling behaviour on average look as if agents generalise
across similar states, even if no agent actually generalises. To explore whether confusion of
states can have this effect, we turn to numerical simulation.

@@

As for utility, we define it in terms of similarity and introduce another free parameter,
$\toler \ge 0$.  The intention is for this parameter to model the amount of tolerable pragmatic
slack, which should be allowed to vary separately from perceptual imprecision:
\begin{align*}
  \util(\mystate{i}, \mystate{j} \myts ; \myts \toler) =
      \similarity(\mystate{i}, \mystate{j} \myts ; \myts \toler)\,.
\end{align*}
This choice of utility function is governed partly by convenience, but also because we believe it has the right general properties for a communicative payoff function.
Unlike utilities that are, say, linearly or
quadratically decreasing in physical distance
(c.f.~\citealtbjps{JagerMetzger2011:Voronoi-Languag}; \citealtbjps{FrankeJager2010:Vagueness-Signa}), utilities
that are exponentially decreasing in negative quadratic distance can model situations where a
small amount of imprecision in communication is tolerable, whereas similarly small differences
in intolerably far away interpretations matter very little, with a smooth transition between
these regimes \citepbjpspre{c.f.}{OConnor2013:The-Evolution-o}.

@@

We ran 50 trials of the discrete-time dynamic in (\ref{eq:nRD}), starting with randomly sampled
sender and receiver strategies, for each triplet of independent parameter values:
$\ns \in \set{6, 10, 50, 90}$, $\imprecision \in \set{0, 0.05, 0.1, 0.2, 0.3}$,
$\toler \in \set{0.05, 0.1, 0.2, 0.3}$. Each trial ran for a maximum of 200 update steps. A
trial was considered converged, and thus stopped before the maximum of 200 rounds, if the total
amount of change between strategies before and after an update step was smaller than a suitably
chosen threshold. It is not guaranteed that strategies at halting time had converged to the
eventual attracting state, whether they ran for 200 rounds or not. Our notion of convergence is
therefore only a categorical measure for reaching a certain (well-considered, but eventually
arbitrary) degree of stability. In other words, our notion of ``convergence'' is a measure of
relative speed: is it true that the system reached a state in which evolutionary adaptations
had slowed down almost to a halt before 200 update steps? This is motivated by pragmatic
concerns regarding length of simulation time, but also theoretically justified, because we
hypothesise that confusability of states leads to regularisation of evolving strategies, which
would show exactly in an increased speed of evolutionary trajectories towards well-behaved and
regular signalling behaviour.

@@

Representative examples for resulting strategy pairs are given in
Figure~\ref{fig:example_strats}. Figure~\ref{fig:example_stratsA} shows a strategy pair at
stopping time with 90 states, tolerance $\toler = 0.1$ and imprecision $\imprecision = 0$. Zero
imprecision means that the trial was effectively an application of the standard replicator dynamic. Noteworthily,
the given sender strategy approximates a pure sender strategy that crisply partitions the state
space into non-convex sets. The irregular shape of the receiver strategy suggests that the
pictured strategy pair has not yet reached a stable state. Indeed,
the trial was stopped when reaching the maximum of 200 rounds. In contrast, the outcome of a
trial with identical parameters, except with imprecision $\imprecision = 0.05$, which is shown
in Figure~\ref{fig:example_stratsB}, had converged (in our technical sense) after 99
rounds. The sender strategy shows a smooth blending from one ``category'' to the other, and the
receiver's interpretations are rather extended curves, peaking at a central point in the
relevant ``categories.''

@@

>>>>

\begin{abstract}
  Signalling games are popular models for studying the evolution of meaning, but typical
  approaches do not incorporate vagueness as a feature of successful signalling.  Complementing
  recent like-minded models, we describe an aggregate population-level dynamic that describes a
  process of imitation of successful behaviour under imprecise perception and realisation of
  similar stimuli. Applying this new dynamic to a generalisation of Lewis' signalling games, we
  show that stochastic imprecision leads to vague, yet by-and-large efficient signal use, and,
  moreover, that it unifies evolutionary outcomes and helps avoid sub-optimal
  categorisation. The upshot of this is that we see `as-if'-generalisation at an aggregate
  level, without agents actually generalising.
\end{abstract}

@@

Formally, a sim-max game consists of a set of states $\States$, a set of messages $\Messgs$
typically with much fewer messages than states, a probability distribution
$\Pr \in \Delta(\States)$ such that $\Pr(\state)$ gives the probability that state $\state$
occurs, a similarity metric\footnote{A metric is a function of distance between any two points
  in a given space. It should satisfy certain axioms that ensure behaviour that one would
  intuitively expect from the words `similarity' and `distance' alone, but these details do
  not matter for the purposes of this paper. The assumption that state similarity forms a
  metric is conceptually loaded, but we follow the literature here and conceive of it as a
  first and pragmatic simplification, possibly to be dispensed with later.}
  % : similarity should always be non-negative, zero only for the distance
  % between an element and itself, symmetric, and satisfy the triangle inequality that
  % $\similarity(x,z) \le \similarity(x,y) + \similarity(y,z)$
on states $\similarity \mycolon \States \times \States \rightarrow \mathds{R}$ such that
$\similarity(\mystate{1}, \mystate{2})$ is the (physical) similarity between $\mystate{1}$ and
$\mystate{2}$, and a utility function
$\Utils \mycolon \States \times \States \rightarrow \mathds{R}$ such that
$\Utils(\mystate{1}, \mystate{2})$ is the payoff for sender and receiver for a play with actual
state $\mystate{1}$ and receiver interpretation $\mystate{2}$. We identify the receiver's acts
with the states of the world, so that the game is one of guessing the actual state, so to
speak. For the modelling purposes of this paper, we make the simplifying assumption that
$\States$ is a set of points in Euclidean space, whose closeness to each other tracks physical
similarity. Perceived similarity, where it is necessary, would be a monotonic function of
physical similarity. Likewise, the utility function should be a monotonically increasing
function of physical similarity.\footnote{Section~\ref{sec:setting-stage} motivates particular
  choices of similarity and utility functions that we will explore in more detail.}
Non-probabilistic sender behaviour can be represented by a pure strategy
$\Spure \in \Messgs^\States$, which deterministically defines which message would be used for
each state. Similarly, a pure receiver strategy is a function $\Rpure \in \States^\Messgs$.

@@

It is intuitive to think that linguistic and conceptual categories are orderly in such a
manner.  For example, if you consider two people `tall', one with a height of 2m and another with a
height of 2.2m, it would be difficult to defend not considering a person with a height
of 2.1m `tall' as well. Another example where this intuition is additionally supported by empirical data is that of colour categorisation. The World Color Survey
project~(\citealtbjps{CookKay2005:The-World-Color}; \citealtbjps{WCS}) collected colour naming data for 110 unwritten
languages of 45 language families.  For the great majority of these languages a pattern can be
observed: basic colour terms are by and large convex~(\citealtbjps{Regier07}; \citealtbjps{Jager2010}). It
is premature to argue that these observations can be extended to all cases of
categorisation. However, for the cases where it does apply, the result of
\citetbjps{JagerMetzger2011:Voronoi-Languag} is interesting because it demonstrates that signalling
can impose this kind of orderly categories on a metric space without that being the ulterior
purpose of it all.%
\footnote{For the concrete case of colour categorisation, see also
  \citetbjps{JagerRooijvan-Rooij2007:Language-Struct} and \citetbjps{Correia2015}}
Finally, these considerations are in line with a prominent
school of thought in comparative linguistics which also assumes that more abstract conceptual domains
(e.g., spatial-topological relations, temporal reference, or the meanings expressible by
indefinite pronouns) are preferably carved up by the languages of the world in such a way
that meanings are connected regions on a so-called `semantic map'
(e.g.~\citealtbjps{Croft2003:Typology-and-Un}; \citealtbjps{Haspelmath2003:The-geometry-of}; \citealtbjps{LevinsonMeira2003:Natural-concept}).

There are at least two potential ways of interpreting the signalling setup.  Sender and
receiver can be distinct entities, whose purpose is to communicate effectively about the actual
state. In that case, evolving Voronoi languages would explain why linguistic categories are
well-behaved and orderly in the way they appear to be. More abstractly, sender and receiver can
also be thought of as distinct modules in a single system, where the first module must
discretise the information it is fed by selecting a small sample of, suggestively, category
labels. These are passed to a second module that tries to decode the original information. In
this case, evolving Voronoi languages would explain why conceptual categories are well-behaved
and orderly in the way that they appear to be.  Seen in this light, sim-max games may provide a
foundation to approaches in cognitive semantics that rely on the notion of conceptual spaces.
\citetbjpspost{pp.~70--77}{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued
that natural categories are convex regions in conceptual space. If the conceptual space has a
suitable metric, convex categories can be derived from a set of prototypes. The category
corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
other. In this way, \citeauthor{Gardenfors2000:Conceptual-Spac} argues, an efficient
categorisation system can be obtained: storing the prototypes lets us recover the categories
without having to store each category's extension. However, what is left unexplained is where
the prototypes come from, and why we would not see just any distribution of prototypes as an
equally efficient classification system. This is where sim-max games can contribute a
principled approach to deriving, in an independent way, not only convex categories but also
prototypical exemplars belonging to them.  These ideas, and more, are developed further in
(\citealtbjps{Jager2007:The-Evolution-o};
\citealtbjps{JagerRooijvan-Rooij2007:Language-Struct};
\citealtbjps{JagerMetzger2011:Voronoi-Languag}; \citealtbjps{OConnor2014-OCOEPC}), among
others.

@@

\citetbjps{DouvenDecock2011:Vagueness:-A-Co} show that
\citeauthor{Gardenfors2000:Conceptual-Spac}'s conceptual spaces
approach can be extended to account for the existence of borderline
cases. From the assumption that prototypes are extended, yet convex
regions in conceptual space, a construction algorithm is available
that yields `collated Voronoi diagrams' with thick boundaries
representing borderline
regions. \citetbjps{DecockDouven2012:What-is-Graded-} show further how it
is possible to arrive at a gradient transition between categories, by
weighing in the distance of different borderline cases to various
prototypical regions. This accounts for the first of the two
desiderata mentioned above, but still assumes that crisp prototype
regions must be given.

@@

Our conceptual starting point is the widely shared conviction that language is shaped by at
least two forces, which may, on occasion, pull in opposite direction. On the one hand, there is
functional pressure towards efficient communication. On the other hand, there is systematic
error, noise or imprecision in the transmission of linguistic behaviour, knowledge or traits. As
an example for the latter, consider a child learning syntactic rules from a parent
generation. The child must infer these unobservable rules from observable speech. Inductive
biases may influence which syntactic rules are likely to be inferred from (finite) parental
input. Over the course of many generations, the effects of such biases can, in a manner of
speaking, `accumulate' and lead to surprising results, such as the evolution of compositional
form-meaning mappings or the use of regular recursive syntactic structure. This can happen when
the effects of transmission biases are isolated, as in iterated learning models
(\citealtbjps{KirbyHurford2002:The-Emergence-o}; \citealtbjps{SmithKirby2003:Iterated-Learni}; \citealtbjps{GriffithsKalish2007:Language-Evolut}; \citealtbjps{KirbyGriffith2014:Iterated-Learni}),
or when they interact with functional pressure towards efficient communication, e.g., as
formalised in the replicator mutator dynamic
(e.g.~\citealtbjps{NowakPlotkin2000:The-Evolution-o}; \citealtbjps{NowakKomarova2001:Evolution-of-Un}).

@@

Fix a sim-max game with finite states $\States$ and messages $\Messgs$. As usual, we assume
that the receiver chooses states in $\States$ in response to messages. Let
$\Pr(\cdot) \in \Delta(\States)$ be the prior distribution over states and
$\Util \mycolon \States \times \States \rightarrow \mathds{R}$ the utility function shared by
senders and receivers in the population. A behavioural strategy is a function that maps
an agent's choice points to a probability distribution over available
choices.\footnote{\label{fn:mixedStrats}Our focus is on behavioural strategies, not mixed
    strategies, i.e., probability distributions over functions from each choice point to an
  act, such as $s \in \Delta(\Messgs^{\States})$.  Dynamics on behavioural strategies assume
  that agents can adapt their behaviour locally, i.e., independently at each choice point. Our
  focus on behavioural strategies greatly reduces the complexity of the dynamic and simplifies
  numerical simulations. But it also seems the more plausible choice for imitation-based update
  protocols of the kind we consider here: agents only observe how, on some occasion, some other
  agent behaved in one particular situation, not how that agent would behave in
  all relevant choice situations; they imitate the use of a single word, so to speak,
  not a whole lexicon. (See \citealtbjps{Cressman2003:Evolutionary-Dy} for more on the difference
  between dynamics on mixed or behavioural strategies.) It is in this respect that the cultural
  evolutionary dynamic introduced here differs most visibly from the replicator mutator dynamic
  (e.g.~\citealtbjps{NowakPlotkin2000:The-Evolution-o}; \citealtbjps{NowakKomarova2001:Evolution-of-Un}), which
  operates on mixed strategies and is motivated by assumptions of (asexual) biological
  inheritance with transmission infidelity.} The sender's behavioural strategies are
functions $\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$
to a probability of each message $\messg \in \Messgs$ being sent in $\state$; the receiver's
are functions $\Rstrat \in \Delta(\States)^\Messgs$, thus mapping each message
$\messg \in \Messgs$ to a probability of each interpretation $\state \in \States$ being chosen
in response to $\messg$. Although behavioural strategies are probabilistic, evolutionary
modelling usually imagines that every individual agent has a non-probabilistic
strategy. Behavioural strategies then capture average population behaviour. Assuming a virtually
infinite population, the number $\Sstrat(\messg \probbar \state)$, for instance, is then the
probability that a randomly sampled sender would send message $\messg$ if the actual state was
$\state$. Similarly, $\Rstrat(\state \probbar \messg)$ is then the probability with which a
randomly sampled receiver interprets $\messg$ as $\state$. A play of a single evolutionary game
with behavioural strategies is illustrated in Figure~\ref{fig:sim-max-illustration}.

@@

The expected utility of choices at each choice point is:\footnote{The notation
  $\propto$, for `proportional to', that is used here and hereafter means that the right-hand
  side might still need to be normalised so as to have probabilities sum to one. In general,
  writing $P(x) \propto f(x)$ for any function $f \mycolon X \rightarrow \mathds{R}$ is
  shorthand for $P(x) = \frac{f(x)}{\sum_{x' \in X} f(x')}$. }
\begin{align*}
  \EU(\messg, \state, \Rstrat) & = \sum_{\state \in \States}
  \Rstrat(\state' \probbar \messg) \myts U(\state, \state') \\
  \EU(\state', \messg, \Sstrat) & = \sum_{\state \in
    \States} P(\state \probbar \messg)  \myts
  U(\state, \state')\,, \text{\ \  where } &
P(\state \probbar \messg) & \propto \Pr(\state) \myts \Sstrat(\messg \probbar \state) \,.
\end{align*}
Expected utilities are sums, over all possible concrete outcomes, of the utilities of these
outcomes, weighted by how likely they occur. For instance, the expected utility for the sender
of choosing message $\messg$, given state $\state$ and a receiver strategy $\Rstrat$ is the
sum, for each interpretation $\state'$, of the probability that some random receiver will
choose $\state'$ given $\messg$ times the actual utility of pair $\state$ and
$\state'$.

The discrete-time replicator dynamic tracks changes in frequency of choices in the population
as proportional to their expected utilities
\citepbjpspre{e.g.}{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The formulation given in
  Equation~(\ref{eq:RDDiscrete}) is adequate only for cases like the one we will be looking at
  in Section~\ref{sec:exploring-rdd}, where utilities are always non-negative and expected
  utilities are always positive.}
\begin{align}
  \label{eq:RDDiscrete}
  \Sstrat'(\messg \probbar \state) & \propto \Sstrat(\messg \probbar \state) \myts
    \EU(\messg, \state,\Rstrat)\,, & \Rstrat'(\state \probbar \messg) & \propto \Rstrat(\state \probbar \messg) \myts
    \EU(\state, \messg,\Sstrat)\,.   
\end{align}
For a given choice point, say a state $\state$, the probability of seeing $\messg$ played by an
average agent in the population after the update is proportional to the probability of
seeing it before the update, which is $\Sstrat(\messg \probbar \state)$, times the
expected utility of $\messg$ at state $\state$. Intuitively put, frequencies of choices change
by a gradient of current frequency and a measure of how good they are.

The replicator dynamic is an abstract population-level dynamic that describes the mean expected
change of behavioural dispositions in a population of signallers. There are several ways of
deriving the replicator dynamic from agent-level processes of behavioural adaptation. We focus
here on one of the simplest: imitation of success
\citepbjpspre{see}{Sandholm2010:Population-Game}. The intuitive idea is the following (see
Appendix~\ref{sec:cond-imit-succ} for a derivation of the standard replicator dynamic from this
update scheme).  Every now and then a random agent gets a chance to alter his behaviour for one
of his choice points. Say, this is a sender who gets to `reconsider' his choice of message
for state $\state$. The revising agent then observes what some random agent does at $\state$,
say $\messg$, and will henceforth play $\messg$ in $\state$ with a probability given by the
expected utility of $\messg$ for state $\state$.

@@

How does a tendency to confuse similar states interact with selective pressure towards more
efficient signalling strategies under the imprecise conditional imitation dynamic? If confusion
probabilities are moderate and regular, in that they track similarity of states, we might
expect a regularising effect also on evolving signalling strategies. Indeed, we hypothesise
that confusion of states can give rise to population-level aggregate behaviour that looks as
if information about what is good for one state percolates also to similar states. In other
words, imprecise imitation may make signalling behaviour on average look as if agents
generalise across similar states, even if no agent actually generalises. To explore whether
confusion of states can have this effect, we turn to numerical simulation.

@@

As for utility, we define it in terms of similarity and introduce another free parameter,
$\toler \ge 0$.  The intention is for this parameter to model the amount of tolerable pragmatic
slack, which should be allowed to vary separately from perceptual imprecision:
\begin{align*}
  \util(\mystate{i}, \mystate{j} \myts ; \myts \toler) =
      \similarity(\mystate{i}, \mystate{j} \myts ; \myts \toler)\,.
\end{align*}
With $\toler = 0$, we regain the case of Lewis' games where only a perfect match of actual
state $\mystate{i}$ with receiver interpretation $\mystate{j}$ leads to a positive payoff. The
higher $\toler$ the more acceptable a wider environment of interpretations $\mystate{j}$ around
$\mystate{i}$ is (see Figure~\ref{fig:NosofskySim}).  This choice of utility function is
governed partly by convenience, but also because we believe it has the right general properties
for a communicative payoff function.  Unlike utilities that are, say, linearly or quadratically
decreasing in physical distance (c.f.~\citealtbjps{JagerMetzger2011:Voronoi-Languag};
\citealtbjps{FrankeJager2010:Vagueness-Signa}), utilities that are exponentially decreasing in
negative quadratic distance can model situations where a small amount of imprecision in
communication is tolerable, whereas similarly small differences in intolerably far away
interpretations matter very little, with a smooth transition between these regimes
\citepbjpspre{c.f.}{OConnor2013:The-Evolution-o}.

@@

We ran 50 trials of the discrete-time dynamic in (\ref{eq:nRD}), starting with randomly sampled
sender and receiver strategies, for each triplet of independent parameter values:
$\ns \in \set{6, 10, 50, 90}$, $\imprecision \in \set{0, 0.05, 0.1, 0.2, 0.3}$,
$\toler \in \set{0.05, 0.1, 0.2, 0.3}$. Each trial ran for a maximum of 200 update steps. A
trial was considered converged, and thus stopped before the maximum of 200 rounds, if the total
amount of change between strategies before and after an update step was smaller than a suitably
chosen threshold. It is not guaranteed that strategies at halting time had converged to the
eventual attracting state, whether they ran for 200 rounds or not. Our notion of convergence is
therefore only a categorical measure for reaching a certain (well-considered, but eventually
arbitrary) degree of stability. In other words, our notion of `convergence' is a measure of
relative speed: is it true that the system reached a state in which evolutionary adaptations
had slowed down almost to a halt before 200 update steps? This is motivated by pragmatic
concerns regarding length of simulation time, but also theoretically justified, because we
hypothesise that confusability of states leads to regularisation of evolving strategies, which
would show exactly in an increased speed of evolutionary trajectories towards well-behaved and
regular signalling behaviour.

@@

Representative examples for resulting strategy pairs are given in
Figure~\ref{fig:example_strats}. Figure~\ref{fig:example_stratsA} shows a strategy pair at
stopping time with 90 states, tolerance $\toler = 0.1$ and imprecision $\imprecision = 0$. Zero
imprecision means that the trial was effectively an application of the standard replicator dynamic. Noteworthily,
the given sender strategy approximates a pure sender strategy that crisply partitions the state
space into non-convex sets. The irregular shape of the receiver strategy suggests that the
pictured strategy pair has not yet reached a stable state. Indeed,
the trial was stopped when reaching the maximum of 200 rounds. In contrast, the outcome of a
trial with identical parameters, except with imprecision $\imprecision = 0.05$, which is shown
in Figure~\ref{fig:example_stratsB}, had converged (in our technical sense) after 99
rounds. The sender strategy shows a smooth blending from one `category' to the other, and the
receiver's interpretations are rather extended curves, peaking at a central point in the
relevant `categories.'

@@

\section*{Acknowledgements}
